
## Loading and preparing the data

The first step is to load and prepare the trade data. This data set was extracted from the Convention on International Trade in Endangered Species of Wild Fauna and Flora (CITES), an international treaty organization tasked with monitoring, reporting, and providing recommendations on the international species trade. 

This data set contains records on every international import or export conducted with species from the CITES lists in 2016. It contains columns identifying the species, the import and export countries, and the amount and characteristics of the goods being traded (which range from live animals to skins and cadavers).

```{r   }

#Loading required libraries 
library(data.table)
library(ggplot2)        
library(dplyr)          
library(igraph)         
library(DT)             
library(shiny)          
library(ggmap)
library(leaflet)
library(bslib)
library(sf)
library(geosphere)
library(RColorBrewer)
library(sp)

# Load data from file cleaned_trade_data.csv 
dt.trade <- fread("../../data/cleaned_trade_data.csv") 

# Save dt.trade 
save(dt.trade, file="trade.RData") 

```

```{r}

#Loading previously saved dt.trade (possible to start in this line if data has been previously saved
load("trade.RData") 

#Creating a list of unique countries SHOULD WE CHANGE THIS TO PARTNER NAME ?????????
l.countries <- as.list(unique(dt.trade$reporter_name))

#Creating a data table with the unique names of countries
dt.country.coordinates <- dt.trade %>% distinct(partner_name, .keep_all = TRUE)

#Selecting only the columns of country name , longitude and latitude and renaming them columns for future plotting
dt.country.coordinates <- dt.country.coordinates[, c("partner_name", "partner_lat", "partner_long")]
meta <- dt.country.coordinates %>% rename("name" = "partner_name", "lat" = "partner_lat", "lon" = "partner_long")

```

```{r}
dt.trade.country <- dt.trade[ dt.trade$reporter_name == "Germany", ]
dt.trade.country.year <- dt.trade.country[dt.trade.country$year >= 2018 &
                                            dt.trade.country$year <= 2021, ]
dt.trade.country.year <- dt.trade.country.year[dt.trade.country.year$trade_value_usd >                                                      0, ]
# Set the vertices
dt.all.reporters <- dt.trade.country.year[, list(name=unique(reporter_name), type=FALSE)]
dt.all.partners <- dt.trade.country.year[, list(name=unique(partner_name), type=TRUE)]

dt.all.vertices <- rbind(dt.all.reporters, dt.all.partners)

# Create & plot the graph
g <- graph.data.frame(dt.trade.country.year[, list(reporter_name, partner_name)], directed=TRUE, vertices=dt.all.vertices)

    
# Setting weights as the trade value 
g <- set_edge_attr(g, "weight", value = dt.trade.country.year$trade_value_usd)

g.subgraph <- subgraph.edges(g, which(V(g)$name == "Germany" ))


df.map <- data.frame(
      DegreeCentrality = degree(g,v="Germany"),
      ClosenessCentrality = closeness(g,v="Germany"),
      BetweennessCentrality = betweenness(g,v="Germany"),
      EigenvectorCentrality = eigen_centrality(g.subgraph)$vector["Germany"],
      ClusteringCoefficient =transitivity(g,v="Germany")
    )
    # Transpose the table
    df.map.t <- t(df.map)
    
    df.map.t


```

````{r}

#SHINY APP WITH MAP


# Define UI --------------------------------------------------------------------

ui <- fluidPage(
  
  # App title ----
  titlePanel("Shiny App for International Trade"),
  
  #Navigation Page to shift from pages 
  navbarPage(title = "Network Analytics",
           tabPanel(title = "Rankings"),
           tabPanel(title = "Descriptive"),
           tabPanel(title = "Interactive Map"),
           theme = bs_theme(bg = "white",
                            fg = "black",
                            primary = "maroon",
                            base_font = font_google("Montserrat")
                            )
          ),
  
  
  sidebarLayout(
    
    # Inputs: Select variables to plot
    sidebarPanel(
      
      #Adding logo to sidebar
      img(src = "https://upload.wikimedia.org/wikipedia/commons/thumb/f/f5/World_Trade_Organization_%28logo_and_wordmark%29.svg/2560px-World_Trade_Organization_%28logo_and_wordmark%29.svg.png",width="100%"),
      
      #Introducing Inputs for the user to select type of trade, country, minimum trade value and years to plot 
      
      # Select Trade
      selectInput(inputId = "trader", 
                  label = "Select Imports or Exports:",
                  choices = c("Exports"="reporter_name","Imports"="partner_name"), 
                  selected = "reporter_name"),
      
      # Select country
      selectInput(inputId = "country", 
                  label = "Select Country:",
                  choices = l.countries, 
                  selected = "Ghana"),
      
      # Select minimum weight
      sliderInput(inputId = "MinWeight",
                label = "Minimum Trade Value (USD):",
                min = 0,
                max = 20000000000,
                value = 0,
                step = 1000000,
                width = "90%"
                ),
      
      # Set year range
      sliderInput(inputId = "year_range",
                  label = "Select Year Range:",
                  min = 2000,
                  max = 2021,
                  value = c(2018, 2021),
                  step = 1)
    ),
    
    
    # Output: Show network
    mainPanel(

      leafletOutput("map"),
      DTOutput("datamap")
    )
  )
)


# Define server ----------------------------------------------------------------

server <- function(input, output, session) {

    #  Centralities table output
  output$datamap <- renderDT({
    
      if(input$trader=="reporter_name"){
        dt.trade.country <- dt.trade[ dt.trade$reporter_name == input$country, ]
        dt.trade.country.year <- dt.trade.country[dt.trade.country$year >= input$year_range[1] &
                                            dt.trade.country$year <= input$year_range[2], ]
        dt.trade.country.year <- dt.trade.country.year[dt.trade.country.year$trade_value_usd >                                                      input$MinWeight, ]
        
        } else {
          dt.trade.country <- dt.trade[ dt.trade$partner_name == input$country, ]
          dt.trade.country.year <- dt.trade.country[dt.trade.country$year >= input$year_range[1] &
                                            dt.trade.country$year <= input$year_range[2], ]
          dt.trade.country.year <- dt.trade.country.year[dt.trade.country.year$trade_value_usd >                                                      input$MinWeight, ]
          
        }
    # Set the vertices
    dt.all.reporters <- dt.trade.country.year[, list(name=unique(reporter_name), type=FALSE)]
    dt.all.partners <- dt.trade.country.year[, list(name=unique(partner_name), type=TRUE)]

    dt.all.vertices <- rbind(dt.all.reporters, dt.all.partners)

    # Create & plot the graph
    g <- graph.data.frame(dt.trade.country.year[, list(reporter_name, partner_name)], directed=TRUE,
                          vertices=dt.all.vertices)

    # Setting weights as the trade value 
    g <- set_edge_attr(g, "weight", value = dt.trade.country.year$trade_value_usd)
    
    g.subgraph <- subgraph.edges(g, which(V(g)$name == input$country))

    # Extract the corresponding row from the data frame

    df.map <- data.frame(
      DegreeCentrality = degree(g, v= input$country),
      ClosenessCentrality = closeness(g, v= input$country),
      BetweennessCentrality = betweenness(g, v= input$country),
      EigenvectorCentrality = eigen_centrality(g.subgraph)$vector[input$country],
      ClusteringCoefficient =transitivity(g, v= input$country)
    )
    # Transpose the table
    df.map.t <- t(df.map)
    
    df.map.t
    }, options = list(searching = FALSE, lengthChange = FALSE))

  output$map <- renderLeaflet({

    # Set the parameters of the inputs to select years, minimum trade, reporter and partner name
    # This is done with an if statement depending if user selected Exports or Imports 
    if(input$trader=="reporter_name"){
      dt.trade.country <- dt.trade[ dt.trade$reporter_name == input$country, ]
      dt.trade.country.year <- dt.trade.country[dt.trade.country$year >= input$year_range[1] &
                                            dt.trade.country$year <= input$year_range[2], ]
      dt.trade.country.year <- dt.trade.country.year[dt.trade.country.year$trade_value_usd >                                                      input$MinWeight, ]

    } else {
      dt.trade.country <- dt.trade[ dt.trade$partner_name == input$country, ]
      dt.trade.country.year <- dt.trade.country[dt.trade.country$year >= input$year_range[1] &
                                            dt.trade.country$year <= input$year_range[2], ]
      dt.trade.country.year <- dt.trade.country.year[dt.trade.country.year$trade_value_usd >                                                      input$MinWeight, ]

    }

    #creating dataframe from the data table and defining from and to parameters
    df <- data.frame("from" = dt.trade.country.year$reporter_name,
                     "to"= dt.trade.country.year$partner_name)

    #checking if there is data from the inputs demanded my the user, if not, map shows a message saying that no data is available

    if (nrow(df) == 0) {
      popup <- paste0("No data available for selected country")
      return(leaflet() %>% addTiles() %>% setView(lng = 0, lat = 0, zoom = 2) %>%
             addPopups(lng = 0, lat = 0, popup = popup))
    }

    # Creating graph with data frame where the vertices are the meta data table containing all the countries with their longitude and latitude as previously explained 
    g <- graph_from_data_frame(df, directed = TRUE, vertices = meta)
    
    # Setting weights as the trade value 
    g <- set_edge_attr(g, "weight", value = dt.trade.country.year$trade_value_usd)

    # Remove vertices with degree 0
    g <- delete.vertices(g, which(degree(g) == 0))

    # Retrieving data frame form the graph to get the coordinates of the vertices by summing longitude and latitude
    gg <- get.data.frame(g, "both")
    vert <- gg$vertices
    coordinates(vert) <- ~lon + lat

    # Creating spatial lines from the edges of the graph with the weight of the trade value as an attribute
    
    # Retrieving the edges of the graph as a data frame from the output of get.data.frame(g, "both").
    edges <- gg$edges
    
    # Retrieving the weights of the edges from the attribute named "weight" in the graph g.
    weights <- E(g)$weight

    # Creating a list of SpatialLines objects from the vertices data frame vert using the edges data frame. For each row i in the edges data frame, it extracts the "from" and "to" vertex names, retrieves the corresponding rows from vert, creates a SpatialLines object from these vertices, and stores it in the list.
    
    edges <- lapply(1:nrow(edges), function(i) {
      as(rbind(vert[vert$name == edges[i, "from"], ],
               vert[vert$name == edges[i, "to"], ]),
         "SpatialLines")
    })

    #This loop assigns unique IDs to the SpatialLines objects in the list edges using the spChFIDs function. The new IDs are character strings generated from the index i.
    
    for (i in seq_along(edges)) {
      edges[[i]] <- spChFIDs(edges[[i]], as.character(i))
    }

    # Combining the SpatialLines objects in the list edges into a single object of class SpatialLines.
    edges <- do.call(rbind, edges)
    
    # Creating a data frame edges_df with two columns: "id" and "weight". The "id" column contains sequential integers, and the "weight" column contains the weights of the edges.
    edges_df <- data.frame(id = seq_along(edges), weight = weights)
    
    # Combining the SpatialLines object edges with the data frame edges_df to create a SpatialLinesDataFrame. The resulting object contains both the spatial information (the lines connecting the vertices) and the attribute information (the weights of the edges)
    edges <- SpatialLinesDataFrame(edges, data = edges_df)

    #defining color palette for weights on the map used below
    pal <- colorNumeric(palette = "YlOrRd", domain = c(min(edges_df$weight), max(edges_df$weight)))
    
    
    leaflet(vert) %>%
      #Format of the map
      addProviderTiles("Stamen.Toner", options = providerTileOptions(noWrap = TRUE, zoomSnap = 0, 
                                                                     attributionControl = FALSE, 
                                                                     backgroundColor = "#f2f2f2",opacity = 
                                                                       0.4)) %>%
      #Adding red circles on edges
      addCircleMarkers(data = vert, radius = 2,
                   color = "red",
                   fillColor = "red",
                   fillOpacity = 0.8,
                   stroke = FALSE) %>%
      
      #Adding lines connecting edges with a color scheme representing the different weights of the trades
      addPolylines(data = edges, weight = 2, color = ~pal(weight)) %>%
      
      #Adding legend for color of the lines representing the weights of exports and import's trade value
      addLegend(position = "bottomright", pal = pal, values = edges_df$weight, 
                title = "Trade Value (USD)", labFormat = labelFormat(suffix = " USD", digits = 0)) 

          })
}


# Create a Shiny app object ----------------------------------------------------

shinyApp(ui = ui, server = server)

```
